COFFEE: an Optimizing Compiler for Finite Element Local Assembly

### 2014-06-24 TACO-paper/coffee.tex
Now really about to break licm-licmap tables
Broke tables; re-wrote impact of licm, ap, op-vect

@@

\begin{table}[h]
\begin{center}
\tbl{Type and variable names used in the various code snippets to identify local assembly objects.}{
\begin{tabulary}{1.0\columnwidth}{C|C|C}
\hline
Object name & Type & Variable name(s) \\\hline
Determinant of the Jacobian matrix & double & det  \\ 
Inverse of the Jacobian matrix & double & K1, K2, ... \\ 
Coordinates & double** & coords\\ 
Fields & double** & w \\ 
Numerical integration weights & double[] & W \\ 
Basis functions (and derivatives) & double[][] & X, Y, X1, ... \\ 
Element matrix & double[][] & A\\ \hline
\end{tabulary}
}
\end{center}
\label{table:map-name-letters}
\end{table}

@@

\begin{table}[t]
\begin{center}
\tbl{Overview of code transformations for Firedrake-generated assembly kernels.}{
\begin{tabulary}{1.0\columnwidth}{C|C}
\hline
Name (Abbreviation) & Parameter \\\hline
Generalized loop-invariant code motion (\emph{licm}) &   \\ 
Padding &  \\ 
Data Alignment & \\ 
Loop interchange      & loops  \\ 
Loop unrolling  & unroll factor \\ 
Outer-product vectorization (\emph{op-vect}) & tile size \\ 
Expression splitting (\emph{split}) & split point, split factor \\ \hline
\end{tabulary}
}
\end{center}
\label{table:code-transformations}
\end{table}

@@

% "vendor compilers such as Intel's" is awkward and unclear. If Intel is the only one you have looked at, then say "the Intel compiler", otherwise name the ones for which you know this to be true.
Since loop unroll and unroll-and-jam of outer loops are fundamental to the exposure of instruction-level parallelism, tuning critical parameters such as the unroll factor is of great importance. However, inspecting assembly code and comparing with other hand-written implementations indicates that recent versions of the Intel compiler are capable of estimating close-to-optimal values for such parameters. This is particularly true for assembly kernels, in which the loop nest is affine, bounds are usually very small and known at compile-time, and memory accesses are stride one. We therefore leave the backend compiler in charge of selecting the unroll factor. This choice also simplifies COFFEE's cost model. The only situation in which we explicitly unroll-and-jam a loop is when \emph{op-vect} is used, since the transformed code seems to prevent the Intel compiler from applying this optimization, even if specific pragmas are added.

@@

In the following, results for the loop order \texttt{ijk} are shown. For the considerations exposed in Section~\ref{sec:pyop2-compiler}, loop interchanges having an inner loop over \texttt{i} caused slow downs; also, interchanging \texttt{j} and \texttt{k} loops while keeping \texttt{i} as outermost loop did not provide any benefits.

\subsection{Impact of Generalized Loop-invariant Code Motion}

\newcommand{\massmatrixtwodresultsnorms}{
\begin{tabulary}{1.0\textwidth}{cccccc|cccc|cccc|cccc}
\cline{3-18}
& & \multicolumn{8}{c}{\texttt{Sandy Bridge}} & \multicolumn{8}{c}{\texttt{Xeon Phi}}\\
\cline{3-18}
& & \multicolumn{4}{c}{\texttt{licm}} & \multicolumn{4}{c}{\texttt{licm-ap}} & \multicolumn{4}{c}{\texttt{licm}} & \multicolumn{4}{c}{\texttt{licm-ap}} \\
\cline{1-18}
& & \texttt{p1} & \texttt{p2} & \texttt{p3} & \texttt{p4} & \texttt{p1} & \texttt{p2} & \texttt{p3} & \texttt{p4} & \texttt{p1} & \texttt{p2} & \texttt{p3} & \texttt{p4} & \texttt{p1} & \texttt{p2} & \texttt{p3} & \texttt{p4} \\[0.1cm]
\texttt{Helmholtz} & \texttt{triangle} & 1.05 & 1.46 & 1.68 & 1.67 & 1.32 & 1.88 & 2.87 & 4.13 & 1,49 & 1,06 & 1,05 & 1,17 & 1,50 & 2,41 & 1,30 & 1,96 \\
\texttt{Helmholtz} & \texttt{tetrahedron} & 1.36 & 2.10 & 2.64 & 2.27 & 1.35 & 3.32 & 2.66 & 3.27 & 1,28 & 1,29 & 2,05 & 1,73 & 1,41 & 1,50 & 2,79 & 2,81 \\
\texttt{Helmholtz} & \texttt{prism} & 2.16 & 2.28 & 2.45 & 2.06 & 2.63 & 2.74 & 2.43 & 2.75 & 1,04 & 2,26 & 1,93 & 1,64 & 2,38 & 2,47 & 2,15 & 1,71\\[0.1cm]
%\cline{1-18}
\texttt{Diffusion} & \texttt{triangle} & 1.09 & 1.68 & 1.97 & 1.64 & 1.38 & 1.99 & 3.07 & 4.28 & 1,07 & 1,06 & 1,18 & 1,16 & 1,08 & 1,88 & 1,20 & 1,97\\
\texttt{Diffusion} & \texttt{tetrahedron} & 1.30 & 2.20 & 3.12 & 2.60 & 1.41 & 3.70 & 3.18 & 3.82 & 1,00 & 1,38 & 2,02 & 1,74 & 1,05 & 1,51 & 2,76 & 3,00\\
\texttt{Diffusion} & \texttt{prism} & 2.15 & 1.82 & 2.71 & 2.32 & 2.55 & 3.13 & 2.73 & 2.69 & 1,11 & 2,16 & 1,85 & 2,83 & 2,41 & 2,52 & 2,05 & 2,48\\[0.1cm]
%\cline{1-18}
\texttt{Burgers} & \texttt{triangle} & 1.53 & 1.81 & 2.68 & 2.46 & 1.56 & 2.28 & 2.61 & 2.77 & 1,21 & 1,42 & 2,34 & 2,97 & 2,84 & 2,26 & 3,96 & 4,27 \\
\texttt{Burgers} & \texttt{tetrahedron} & 1.61 & 2.24 & 1.69 & 1.59 & 1.61 & 2.10 & 1.60 & 1.78 & 1,01 & 2,55 & 0,98 & 1,21 & 1,48 & 3,83 & 1,55 & 1,29 \\
\texttt{Burgers} & \texttt{prism} & 2.11 & 2.20 & 1.66 & 1.32 & 2.19 & 2.32 & 1.64 & 1.42 & 1,39 & 1,56 & 1,18 & 1,04 & 2,18 & 2,82 & 1,24 & 1,25 \\
\cline{1-18}
\end{tabulary}
}

\begin{table*}[t]
\tbl{Performance improvement due to generalized loop-invariant code motion (\emph{licm} columns) on the Helmholtz, Diffusion and Burgers problems, varying the element shape (triangle, tetrahedron, prism) and the polynomial order ($p \in [1, 4]$), using the Lagrange element family, over the original, non-optimized code. The \emph{licm-ap} columns illustrate the effect of combining \emph{licm} with data alignment and padding. Results are shown for both the Sandy Bridge and the Xeon Phi architectures.}{
\scriptsize
\massmatrixtwodresultsnorms
}
\label{table:perf-results-licm}
\end{table*}
Table~\ref{table:perf-results-licm} illustrates the performance improvement obtained when \emph{licm} is used. The gain grows with the computational cost of the kernels: bigger loop nests usually benefit from the reduction in redundant computation, even though extra memory for the temporary arrays is required. Inspection of assembly code generated by \texttt{icc} confirmed all limitations described in Section~\ref{sec:code-transf-licm}: in the original version of the code, only sub-expressions invariant to the inner loop were hoisted, and their execution was not vectorized, which motivates the generally good speed ups achieved by \emph{licm}.

\subsection{Impact of Padding and Data Alignment}
Table~\ref{table:perf-results-licmap} shows the cumulative impact of \emph{licm}, data alignment, and padding over the original code. In the following, this version of the code is referred to as \emph{licm-ap}. Padding and data alignment enhance, in general, the quality of auto-vectorization. Occasionally the run-time of \emph{licm-ap} is close to that of \emph{licm}, since the element matrix size is already a multiple of the vector length, and no penalty for a scalar reminder loop has to be paid. Rarely \emph{licm-ap} is slower than \emph{licm} (e.g. in Burgers $p=3$ on the Sandy Bridge). One possible explanation is that the number of aligned temporaries introduced by \emph{licm} is so large to induce cache associativity conflicts.

\subsection{Impact of Vector-register Tiling}
\label{sec:perf-eval-opvect}
Figures~\ref{fig:opvect-helmholtz-speedup} and~\ref{fig:opvect-diffusion-speedup} show the performance achieved applying \emph{op-vect} on top of \emph{licm-ap} to the Helmoltz and Diffusion kernels on the Sandy Bridge and Xeon Phi architectures, respectively. For each problem instance we report two bars: one indicates the best run-time obtained by auto-tuning the unroll/unroll-and-jam factors, whereas the other shows the result retrieved with COFFEE's cost model. In general, there is no substantial difference between the two. This is chiefly because these kernels fit the L1 cache, so, within a certain degree of confidence, it is possible to predict the fastest implementation based on register allocation.

The rationale behind these results is that, for smaller configurations, vector-register tiling is marginally helpful, while its impact is the dominant factor for larger problems. This is because in the considered kernels the number of accessed arrays along the innermost loop dimension is rather small (between 2 and 4), so extensive unrolling is often enough to maximize register re-use when the loops are relatively short. On the other hand, as the iteration space becomes bigger, vector-register tiling leads to improvements up to 1.4$\times$ on the Sandy Bridge (Diffusion, prismatic mesh, $p=4$) and up to 1.4$\times$ on the Xeon Phi (Helmholtz, tetrahedral mesh, $p=3$). Using the Intel Architecture Code Analyzer tool~\cite{IACA} on the Sandy Bridge, we confirmed that this is a consequence of increased register re-use. In Helmholtz $p=4$, for example, the tool showed that when using \emph{op-vect} the number of clock cycles to execute one iteration of the \texttt{j} loop decreases by roughly 17$\%$, and that this is a result of the relieved pressure on both of the data (cache) ports available in the core.

@@

\begin{figure}[h]
	\centering
	\subfloat[Sandy Bridge]{	\includegraphics[scale=0.52]{Pictures/helmholtz-normalized-opvect.pdf}}
	\subfloat[Xeon Phi]{	\includegraphics[scale=0.52]{Pictures/helmholtz-normalized-opvect-phi.pdf}}
	\caption{Performance improvement obtained applying \emph{op-vect} on top of \emph{licm-ap} to the Helmholtz kernel. CM refers to where the cost model has been used to determine the sequence of code transformations.}
	\label{fig:opvect-helmholtz-speedup}
\end{figure}

\begin{figure}[h]
	\centering
	\subfloat[Sandy Bridge]{	\includegraphics[scale=0.52]{Pictures/diffusion-normalized-opvect.pdf}}
	\subfloat[Xeon Phi]{	\includegraphics[scale=0.52]{Pictures/diffusion-normalized-opvect-phi.pdf}}
	\caption{Performance improvement obtained applying \emph{op-vect} on top of \emph{licm-ap} to the Diffusion kernel. CM refers to where the cost model has been used to determine the sequence of code transformations.}
	\label{fig:opvect-diffusion-speedup}
\end{figure}

@@

\begin{figure}[h]
	\centering
	\subfloat[Sandy Bridge]{	\includegraphics[scale=0.52]{Pictures/burgers-normalized-split.pdf}}
	\subfloat[Xeon Phi]{	\includegraphics[scale=0.52]{Pictures/burgers-normalized-split-phi.pdf}}
	\caption{Performance improvement obtained applying \emph{split} on top of \emph{licm-ap} to the Burgers kernel. CM refers to where the cost model has been used to determine the sequence of code transformations.}
	\label{fig:split-burgers-speedup}
\end{figure}

@@

\begin{table}[h]
\begin{center}
\tbl{Performance comparison between FEniCS (optimizations enabled) and COFFEE on the Sandy Bridge}{
\label{table:comparison-to-FFC-opt}
\begin{tabulary}{1.0\columnwidth}{C|C|C}
\hline
Problem & Max slow down & Max speed up \\\hline
Helmholtz & - & 4.14$\times$ \\
Diffusion & - & 4.28$\times$ \\
Burgers & 2.24$\times$ & 1.61$\times$ \\\hline
\end{tabulary}
}
\end{center}
\end{table}

>>>>

\begin{table}[b]
\tbl{Type and variable names used in the various listings to identify local assembly objects.}{
\begin{tabulary}{1.0\columnwidth}{C|C|C}
\hline
Object name & Type & Variable name(s) \\\hline
Determinant of the Jacobian matrix & double & det  \\ 
Inverse of the Jacobian matrix & double & K1, K2, ... \\ 
Coordinates & double** & coords\\ 
Fields & double** & w \\ 
Numerical integration weights & double[] & W \\ 
Basis functions (and derivatives) & double[][] & X, Y, X1, ... \\ 
Element matrix & double[][] & A\\ \hline
\end{tabulary}
}
\label{table:map-name-letters}
\end{table}

@@

@@

\begin{table}[t]
\tbl{Overview of code transformations for Firedrake-generated assembly kernels.}{
\begin{tabulary}{1.0\columnwidth}{C|C}
\hline
Name (Abbreviation) & Parameter \\\hline
Generalized loop-invariant code motion (\emph{licm}) &   \\ 
Padding &  \\ 
Data Alignment & \\ 
Loop interchange      & loops  \\ 
Loop unrolling  & unroll factor \\ 
Outer-product vectorization (\emph{op-vect}) & tile size \\ 
Expression splitting (\emph{split}) & split point, split factor \\ \hline
\end{tabulary}
}
\label{table:code-transformations}
\end{table}

@@

% "vendor compilers such as Intel's" is awkward and unclear. If Intel is the only one you have looked at, then say "the Intel compiler", otherwise name the ones for which you know this to be true.
Since loop unroll and unroll-and-jam of outer loops are fundamental to the exposure of instruction-level parallelism, tuning critical parameters such as the unroll factor is of great importance. However, inspecting assembly code and comparing with other hand-written implementations indicates that recent versions of the Intel compiler are capable of estimating close-to-optimal values for such parameters. This is particularly true for assembly kernels, in which the loop nest is affine, bounds are usually very small and known at compile-time, and memory accesses are stride one. We therefore leave the backend compiler in charge of selecting the unroll factor. This choice also simplifies the COFFEE's cost model. The only situation in which we explicitly unroll-and-jam a loop is when \emph{op-vect} is used, since the transformed code seems to prevent the Intel compiler from applying this optimization, even if specific pragmas are added.

@@

\subsection{Loop permutation}
In the following, only results for the loop order \texttt{ijk} are shown. For the considerations exposed in Section~\ref{sec:pyop2-compiler}, loop interchanges having an inner loop along \texttt{i} caused slow downs; also, interchanging \texttt{j} and \texttt{k} loops while keeping \texttt{i} as outermost loop did not provide any benefits.

\subsection{Impact of Generalized Loop-invariant Code Motion}
\label{licm-impact}

\newcommand{\licmresultsnorms}{
\begin{tabulary}{1.0\textwidth}{cccccc|cccc}
\cline{3-10}
& & \multicolumn{4}{c}{\texttt{Sandy Bridge}} & \multicolumn{4}{c}{\texttt{Xeon Phi}} \\
\cline{1-10}
\texttt{problem} & \texttt{shape} & \texttt{p1} & \texttt{p2} & \texttt{p3} & \texttt{p4} & \texttt{p1} & \texttt{p2} & \texttt{p3} & \texttt{p4} \\
\texttt{Helmholtz} & \texttt{triangle} & 1.05 & 1.46 & 1.68 & 1.67 & 1,49 & 1,06 & 1,05 & 1,17 \\
\texttt{Helmholtz} & \texttt{tetrahedron} & 1.36 & 2.10 & 2.64 & 2.27 & 1,28 & 1,29 & 2,05 & 1,73 \\
\texttt{Helmholtz} & \texttt{prism} & 2.16 & 2.28 & 2.45 & 2.06 & 1,04 & 2,26 & 1,93 & 1,64 \\[0.1cm]
%\cline{1-18}
\texttt{Diffusion} & \texttt{triangle} & 1.09 & 1.68 & 1.97 & 1.64 & 1,07 & 1,06 & 1,18 & 1,16 \\
\texttt{Diffusion} & \texttt{tetrahedron} & 1.30 & 2.20 & 3.12 & 2.60 & 1,00 & 1,38 & 2,02 & 1,74\\
\texttt{Diffusion} & \texttt{prism} & 2.15 & 1.82 & 2.71 & 2.32 & 1,11 & 2,16 & 1,85 & 2,83\\[0.1cm]
%\cline{1-18}
\texttt{Burgers} & \texttt{triangle} & 1.53 & 1.81 & 2.68 & 2.46 & 1,21 & 1,42 & 2,34 & 2,97  \\
\texttt{Burgers} & \texttt{tetrahedron} & 1.61 & 2.24 & 1.69 & 1.59 & 1,01 & 2,55 & 0,98 & 1,21  \\
\texttt{Burgers} & \texttt{prism} & 2.11 & 2.20 & 1.66 & 1.32 & 1,39 & 1,56 & 1,18 & 1,04 \\
\cline{1-10}
\end{tabulary}
}

\begin{table*}[t]
\tbl{Performance improvement due to generalized loop-invariant code motion, for different element shapes (triangle, tetrahedron, prism) and polynomial orders ($p \in [1, 4]$), over the original non-optimized code, for the Helmholtz, Diffusion and Burgers problems.}{
\scriptsize
\licmresultsnorms
}
\label{table:perf-results-licm}
\end{table*}
Table~\ref{table:perf-results-licm} illustrates the performance improvement obtained when \emph{licm} is applied. In general, the speed ups are notable. The main reasons were anticipated in Section~\ref{sec:code-transf-licm}: in the original code, 1) sub-expressions invariant to outer loops are not automatically hoisted, while 2) sub-expressions invariant to the innermost loop are hoisted, but their execution is not auto-vectorized. These observations come from inspection of assembly code generated by the compiler.

The gain tends to grow with the computational cost of the kernels: bigger loop nests (i.e. larger element shapes and polynomial orders) usually benefit from the reduction in redundant computation, even though extra memory for the temporary arrays is required. Some discrepancies to this trend are due to a less effective auto-vectorization. For instance, on the Sandy Bridge, the improvement at $p=3$ is larger than that at $p=4$ because, in the latter case, the size of the innermost loop is not a multiple of the vector length, and a reminder scalar loop is introduced at compile time. Since the loop nest is small, the cost of executing the extra scalar iterations can have a significant impact. The reminder loop overhead is more pronounced on the Xeon Phi, where the vector length is twice as big, which leads to proportionally larger scalar reminder loops.

\subsection{Impact of Padding and Data Alignment}

\newcommand{\licmapresultsnorms}{
\begin{tabulary}{1.0\textwidth}{cccccc|cccc}
\cline{3-10}
& & \multicolumn{4}{c}{\texttt{Sandy Bridge}} & \multicolumn{4}{c}{\texttt{Xeon Phi}} \\
\cline{1-10}
\texttt{problem} & \texttt{shape} & \texttt{p1} & \texttt{p2} & \texttt{p3} & \texttt{p4} & \texttt{p1} & \texttt{p2} & \texttt{p3} & \texttt{p4} \\[0.1cm]
\texttt{Helmholtz} & \texttt{triangle} & 1.32 & 1.88 & 2.87 & 4.13 & 1,50 & 2,41 & 1,30 & 1,96 \\
\texttt{Helmholtz} & \texttt{tetrahedron} & 1.35 & 3.32 & 2.66 & 3.27 & 1,41 & 1,50 & 2,79 & 2,81 \\
\texttt{Helmholtz} & \texttt{prism} & 2.63 & 2.74 & 2.43 & 2.75 & 2,38 & 2,47 & 2,15 & 1,71\\[0.1cm]
%\cline{1-18}
\texttt{Diffusion} & \texttt{triangle} & 1.38 & 1.99 & 3.07 & 4.28 & 1,08 & 1,88 & 1,20 & 1,97\\
\texttt{Diffusion} & \texttt{tetrahedron} & 1.41 & 3.70 & 3.18 & 3.82 & 1,05 & 1,51 & 2,76 & 3,00\\
\texttt{Diffusion} & \texttt{prism} & 2.55 & 3.13 & 2.73 & 2.69 & 2,41 & 2,52 & 2,05 & 2,48\\[0.1cm]
%\cline{1-18}
\texttt{Burgers} & \texttt{triangle} & 1.56 & 2.28 & 2.61 & 2.77 & 2,84 & 2,26 & 3,96 & 4,27 \\
\texttt{Burgers} & \texttt{tetrahedron} & 1.61 & 2.10 & 1.60 & 1.78 & 1,48 & 3,83 & 1,55 & 1,29 \\
\texttt{Burgers} & \texttt{prism} & 2.19 & 2.32 & 1.64 & 1.42 & 2,18 & 2,82 & 1,24 & 1,25 \\
\cline{1-10}
\end{tabulary}
}

\begin{table*}[t]
\tbl{Performance improvement due to generalized loop-invariant code motion, data alignment, and padding, for different element shapes (triangle, tetrahedron, prism) and polynomial orders ($p \in [1, 4]$), over the original non-optimized code, for the Helmholtz, Diffusion and Burgers problems.}{
\scriptsize
\licmapresultsnorms
}
\label{table:perf-results-licmap}
\end{table*}

Table~\ref{table:perf-results-licmap} shows the cumulative impact of \emph{licm}, data alignment, and padding over the original code. In the following, this version of the code is referred to as \emph{licm-ap}. Padding, which avoids the introduction of a reminder loop as described in Section~\ref{licm-impact}, as well as data alignment, enhance the quality of auto-vectorization. Occasionally the run-time of \emph{licm-ap} is close to that of \emph{licm}, since the non-padded element matrix size is already a multiple of the vector length. Rarely \emph{licm-ap} is slower than \emph{licm} (e.g. in Burgers $p=3$ on the Sandy Bridge). One possible explanation is that the number of aligned temporaries introduced by \emph{licm} is so large to induce cache associativity conflicts.

\subsection{Impact of Vector-register Tiling}
\label{sec:perf-eval-opvect}
In this section, we evaluate the impact of vector-register tiling. We compare two versions: the baseline, \emph{licm-ap}; and vector-register tiling on top of \emph{licm-ap}, which in the following is referred to simply as \emph{op-vect}.

Figures~\ref{fig:opvect-helmholtz-speedup} and~\ref{fig:opvect-diffusion-speedup} illustrate the speed up achieved by \emph{op-vect} over \emph{licm-ap} in the Helmoltz and Diffusion kernels, respectively. As explained in Section~\ref{sec:pyop2-compiler}, the \emph{op-vect} version requires the unroll-and-jam factor to be explicitly set. To distinguish between the two ways this parameter was determined, for each problem instance (equation, element shape, polynomial order) we report two bars: one shows the best speed-up obtained after all feasible unroll-and-jam factors were tried; the other shows the speed up when the unroll-and-jam factor was retrieved via the cost model. In a plot legend, cost model bars are suffixed with ``CM''.

It is worth noticing that, in most cases, the cost model successfully determines how to transform a kernel to maximize its performance. This is chiefly because assembly kernels fit the L1 cache, so, within a certain degree of confidence, it is possible to predict how to obtain a fast implementation by simply reasoning on the register pressure. For each problem, the cost model stated to use the default loop permutation, to apply a particular unroll-and-jam factor, and not to perform expression splitting, which, as explained in Section~\ref{sec:perf-results-split}, only deteriorates performance in Helmholtz and Diffusion.

The rationale behind these results is that the effect of \emph{op-vect} is significant in problems in which the assembly loop nest is relatively big. When the loops are short, since the number of arrays accessed at every loop iteration is rather small (between 5 and 9), there is no need for vector-register tiling; extensive unrolling is sufficient to improve register re-use and, therefore, to maximize the performance. However, as the iteration space becomes larger, \emph{op-vect} leads to improvements up to 1.4$\times$ on the Sandy Bridge (Diffusion, prismatic mesh, $p=4$ - increasing the overall speed up from 2.69$\times$ to 3.87$\times$), and up to 1.4$\times$ on the Xeon Phi (Helmholtz, tetrahedral mesh, $p=3$ - bringing the overall speed up from 1.71$\times$ to 2.42$\times$).

Using the Intel Architecture Code Analyzer tool~\cite{IACA} on the Sandy Bridge, we confirmed that speed ups are a consequence of increased register re-use. In Helmholtz $p=4$, for example, the tool showed that when using \emph{op-vect} the number of clock cycles to execute one iteration of the \texttt{j} loop decreases by roughly 17$\%$, and that this is a result of the relieved pressure on both of the data (cache) ports available in the core.

@@

\begin{figure}[h]
	\centering
	\subfloat[Sandy Bridge]{	\includegraphics[scale=0.52]{Pictures/helmholtz-normalized-opvect.pdf}}
	\subfloat[Xeon Phi]{	\includegraphics[scale=0.52]{Pictures/helmholtz-normalized-opvect-phi.pdf}}
	\caption{Performance improvement over \emph{licm-ap} obtained by \emph{op-vect} in the Helmholtz kernel. Bars suffixed with ``CM'' indicate that the cost model was used to transform the kernel.}
	\label{fig:opvect-helmholtz-speedup}
\end{figure}

\begin{figure}[h]
	\centering
	\subfloat[Sandy Bridge]{	\includegraphics[scale=0.52]{Pictures/diffusion-normalized-opvect.pdf}}
	\subfloat[Xeon Phi]{	\includegraphics[scale=0.52]{Pictures/diffusion-normalized-opvect-phi.pdf}}
	\caption{Performance improvement over \emph{licm-ap} obtained by \emph{op-vect} in the Diffusion kernel. Bars suffixed with ``CM'' indicate that the cost model was used to transform the kernel.}
	\label{fig:opvect-diffusion-speedup}
\end{figure}

@@

\begin{figure}[h]
	\centering
	\subfloat[Sandy Bridge]{	\includegraphics[scale=0.52]{Pictures/burgers-normalized-split.pdf}}
	\subfloat[Xeon Phi]{	\includegraphics[scale=0.52]{Pictures/burgers-normalized-split-phi.pdf}}
	\caption{Performance improvement over \emph{licm-ap} obtained by \emph{split} in the Burgers kernel. Bars suffixed with ``CM'' indicate that the cost model was used to transform the kernel.}
	\label{fig:split-burgers-speedup}
\end{figure}

@@

\begin{table}[h]
\tbl{Performance comparison between FEniCS (optimizations enabled) and COFFEE on the Sandy Bridge}{
\begin{tabulary}{1.0\columnwidth}{C|C|C}
\hline
Problem & Max slow down & Max speed up \\\hline
Helmholtz & - & 4.14$\times$ \\
Diffusion & - & 4.28$\times$ \\
Burgers & 2.24$\times$ & 1.61$\times$ \\\hline
\end{tabulary}
}
\label{table:comparison-to-FFC-opt}
\end{table}
